---
output:
  github_document:
    md_extensions: "-smart"
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rxode2lincmt

`rxode2lincmt` provides the analytic one, two and three compartment linear
pharmacokinetic solutions and their parameter gradients that
[rxode2](https://nlmixr2.github.io/rxode2/)'s `linCmt()` uses.  The gradients
come from 'stan' automatic differentiation.

This code used to be compiled inside 'rxode2'.  It lives in its own package so
that installing 'rxode2' does not compile 'stan': 'rxode2' imports
`rxode2lincmt` and calls the compiled kernels through an external-pointer
table.  Most users never call `rxode2lincmt` directly; they write `linCmt()` in
an 'rxode2' or 'nlmixr2' model.

The closed-form solutions follow the idea of the
[wnl](https://CRAN.R-project.org/package=wnl) package by Kyun-Seop Bae, though
the implementation here is different.

## Installation

You can install the development version of `rxode2lincmt` from
[GitHub](https://github.com/nlmixr2/rxode2lincmt) with:

``` r
# install.packages("devtools")
devtools::install_github("nlmixr2/rxode2lincmt")
```

Compiling the 'stan'-based solutions takes a while.

## Examples

The eigenvalues and coefficient matrices of a two compartment model from its
micro-constants:

```{r example-solcomp}
library(rxode2lincmt)
solComp2(k10 = 0.1, k12 = 3, k21 = 1)
```

The micro-constants of a two compartment model given clearance and volume
parameters:

```{r example-micros}
linCmtMicros(p1 = 2, v1 = 20, p2 = 3, p3 = 40, ncmt = 2)
```

One step of the per-row kernel: the concentration one time unit after a
100-unit bolus into a one compartment model with clearance 2 and volume 20,
which matches the closed form `100 / 20 * exp(-2 / 20)`:

```{r example-kernel}
step <- linCmtModelDouble(dt = 1, p1 = 2, v1 = 20, p2 = 0, p3 = 0, p4 = 0, p5 = 0,
                          ka = 0, alastNV = 100, rateNV = 0, ncmt = 1L, oral0 = 0L,
                          trans = 1L, deriv = FALSE, type = 0L, tau = 0, tinf = 0,
                          amt = 0, bolusCmt = 0L, ndiff = 0L)
step$val
100 / 20 * exp(-2 / 20)
```
